\chapter{Introduction}

The space of computations of a program may have a complex structure.
Consider a side-effect free programming language.
The possible computations of a tuple $(A,B)$
are rewrite sequences $(A,B) \to \hdots \to (A',B')$.
These sequences can always be decomposed as two non-interfering computations
$A \to \hdots \to A'$ and $B \to \hdots \to B'$.
The reason is that the subexpressions $A$ and $B$ cannot interact with each other.
Indeed, the space of computations of $(A,B)$ can be understood as
the {\em product} of the spaces of $A$ and $B$.
In contrast, the space of computations of a function application $f(A)$ is not so easy
to characterize. The difficulty is that $f$ may use the value of $A$ zero, one, or
possibly many times.

The motivation of this paper is attempting to {\bf understand spaces of computations}.
We hope that this may be helpful to better understand the properties of {\em evaluation strategies},
such as call-by-name or call-by-value,
from a quantitative point of view.
A better understanding may also suggest program optimizations,
and it should allow to justify that certain program conversions are sound:
\eg that they do not turn a terminating program into a non-terminating one.

The pure $\lambda$-calculus is the quintessential functional programming language.
Computations in the $\lambda$-calculus have been thoroughly
studied since its conception in the 1930s.
The well-known theorem by Church and Rosser~\cite{church1936some}
states that the $\lambda$-calculus is \emph{confluent},
which means, in particular, that terminating programs have unique normal forms.
Another result by Curry and Feys~\cite{curry1958combinatory}
states that computations in the
$\lambda$-calculus may be \emph{standardized},
meaning that they may be converted into a computation in canonical form.
A refinement of this theorem by L\'evy~\cite{Tesis:Levy:1978}
asserts that the canonical computation thus obtained is equivalent to the
original one in a strong sense, namely that they are \emph{permutation equivalent}.
In a series of papers~\cite{DBLP:conf/ctcs/Mellies97,DBLP:journals/logcom/Mellies00,DBLP:conf/rta/Mellies02,mellies2002axiomatic,DBLP:conf/birthday/Mellies05},
Melli\`es generalized many of these results to the abstract setting of {\em axiomatic rewrite systems}.

Let us discuss ``spaces of computations'' more precisely.
The \emph{derivation space} of an object~$x$ in some rewriting system
is the set of all {\em derivations}, \ie sequences of rewrite steps, starting from $x$.
In this paper, our rewriting system of interest will be the pure $\lambda$-calculus,
and we will be interested in \emph{finite} derivations only.
In the $\lambda$-calculus,
a transitive relation between derivations may be defined, the {\em prefix order}.
A derivation $\redseq$ is a prefix of a derivation $\redseqtwo$,
written $\redseq \permle \redseqtwo$,
whenever $\redseq$ performs less computational work than $\redseqtwo$.
Formally, $\redseq \permle \redseqtwo$ is defined to hold whenever the {\em projection} $\redseq/\redseqtwo$ 
is empty\footnote{The notion of projection defined by means of residuals
is the standard one, see \eg \cite[Chapter~12]{Barendregt:1984}
or \cite[Section~8.7]{Terese}.}.
For example, if $K = \lam{\var}{\lam{\vartwo}{\var}}$,
the derivation space of the term $(\lam{\var}{\var\var})(K\varthree)$
can be depicted with the \emph{reduction graph} below.
Derivations are directed paths in the reduction graph,
and $\redseq$ is a prefix of $\redseqtwo$ if there is a directed path from the
target of $\redseq$ to the target of $\redseqtwo$.
For instance, $\redextwo\redex_2$ is a prefix of $\redex\redextwo'\redexthree'$:
\[
  \xymatrix@R=.1cm{
    (\lam{\var}{\var\var})(K\varthree)
      \ar^-{\redex}[rr]
      \ar@/_.2cm/^-{\redextwo}[rdd]
  &
  &
    (\lam{\var}{\var\var})(\lam{\vartwo}{\varthree})
    \ar@/^.3cm/[rd]^{\redextwo'}
  &
  \\
  &
  &
    (K\varthree)(\lam{\vartwo}{\varthree}) \ar[r]^-{\redex'_2}
  &
    (\lam{\vartwo}{\varthree})(\lam{\vartwo}{\varthree})
    \ar@/^.2cm/[rd]^-{\redexthree'}
  \\
  &
    (K\varthree)(K\varthree) \ar@/^.3cm/[ru]^{\redex_1} \ar[r]^-{\redex_2}
  &
    (\lam{\vartwo}{\varthree})(K\varthree)
      \ar@/_.2cm/[ru]^-{\redex'_1}
      \ar[rr]^-{\redexthree}
  &
  &
    \varthree
  }
\]
Remark that the relation $\permle$ is reflexive and transitive but not antisymmetric,
\ie it is a quasi-order but not an order.
For example $\redex\redextwo' \permle \redextwo\redex_1\redex_2' \permle \redex\redextwo'$
but $\redex\redextwo' \neq  \redextwo\redex_1\redex_2'$.
Antisymmetry may be recovered as usual when in presence of a quasi-order,
namely by working modulo {\em permutation equivalence}:
two derivations $\redseq$ and $\redseqtwo$
are said to be permutation equivalent,
written $\redseq \permeq \redseqtwo$, if $\redseq \permle \redseqtwo$
and $\redseqtwo \permle \redseq$.
Working modulo permutation equivalence is reasonable because L\'evy's formulation
of the standardization theorem ensures that permutation equivalence is decidable,
and each equivalence class has a canonical representative.

Derivation spaces are known to exhibit various
regularities~\cite{Tesis:Levy:1978,DBLP:journals/tcs/Zilli84,laneve1994distributive,thesismellies,levy_redex_stability,DBLP:conf/lics/AspertiL13}.
In his PhD thesis, L\'evy~\cite{Tesis:Levy:1978} showed that the derivation space of a term is
an upper semilattice: any two derivations $\redseq,\redseqtwo$ from a term $\tm$
have a {\em least upper bound} $\redseq \sqcup \redseqtwo$, defined as $\redseq(\redseqtwo/\redseq)$,
unique up to permutation equivalence.
On the other hand, the derivation space of a term $\tm$ is not an easy structure to understand in general\footnote{Problem~2 in
the RTA List of Open Problems~\cite{dershowitz1991open} poses
the open-ended question of investigating the properties of ``spectra'', \ie derivation spaces.}.
For example, relating the derivation space of an application $\tm\tmtwo$
with the derivation spaces of $\tm$ and $\tmtwo$ appears to be a hard problem.
L\'evy also noted that the {\em greatest lower bound} of two derivations does
not necessarily exist, meaning that the derivation space of a term does not form a lattice in general.
Even when it forms a lattice, it may not necessarily be a {\em distributive} lattice, as observed for example by
Laneve~\cite{laneve1994distributive}.

Consider the following counterexample,
due to L\'evy, showing that the meet of derivations is not well-defined.
Let $\Omega = (\lam{\var}{\var\var})\lam{\var}{\var\var}$,
and consider the reduction space of $(\lam{\var}{a})((\lam{\var}{b})\Omega)$,
where the steps $T_i$ contract $\Omega$:
\[
  \xymatrix{
    &
    (\lam{\var}{a})((\lam{\var}{b})\Omega)
    \ar_{T_1}[d]
    \ar@/_1cm/_{R}[dddl]
    \ar@/^1cm/^{S}[dddr]
    &
  \\
    &
    (\lam{\var}{a})((\lam{\var}{b})\Omega)
    \ar@/_1cm/[ddl] \ar@/^1cm/[ddr]
    \ar_{T_2}[d]
    &
  \\
    &
    (\lam{\var}{a})((\lam{\var}{b})\Omega)
    \ar@/_1cm/[dl] \ar@/^1cm/[dr]
    \ar_{T_3}[d]
    &
  \\
    a
    &
    \vdots
    &
    (\lam{\var}{a})b
  }
\]
Observe that for all $n \in \mathbb{N}$ we have that $T_1 \hdots T_n \sqsubseteq R$
and $T_1 \hdots T_n \sqsubseteq S$ so the meet $R \sqcap S$ does not exist.


In \cite{DBLP:conf/ctcs/Mellies97}, Melli\`es showed that
derivation spaces in any rewriting system satisfying certain axioms
may be factorized using two spaces,
one of {\em external} and one of {\em internal} derivations.



The difficulty to understand derivation spaces is due to three pervasive phenomena
of \emph{interaction} between computations.
The first phenomenon is \emph{duplication}:
in the reduction graph of above, the step $\redextwo$ duplicates the step $\redex$,
resulting in two copies of $\redex$: the steps $\redex_1$ and $\redex_2$.
In such situation, one says that $\redex_1$ and $\redex_2$ are \emph{residuals} of $\redex$,
and, conversely, $\redex$ is an \emph{ancestor} of $\redex_1$ and $\redex_2$.
The second phenomenon is \emph{erasure}:
in the diagram above, the step $\redexthree$ erases the step $\redex'_1$,
resulting in no copies of $\redex'_1$.
The third phenomenon is \emph{creation}:
in the diagram above, the step $\redex_2$ creates the step $\redexthree$,
meaning that $\redexthree$ is not a residual of a step that existed prior
to executing $\redex_2$; that is, $\redexthree$ has no ancestor.

These three interaction phenomena, especially duplication and erasure,
are intimately related with the management of \emph{resources}.
In this work, we aim to explore the hypothesis that {\bf having an explicit
representation of resource management may provide insight on
the structure of derivation spaces}.

There are many existing $\lambda$-calculi that deal with resource management explicitly~\cite{boudol1993lambda,ehrhard2003differential,kesner2007resource,kesner2009prismoid},
most of which draw inspiration from Girard's Linear Logic~\cite{girard1987linear}.
In recent years, one family of such formalisms, namely calculi endowed with
{\em non-idempotent intersection type systems},
has received some attention~\cite{ehrhard2012collapsing,bernadet2013non,bucciarelli2014inhabitation,bucciarelli2017non,kesner2016reasoning,thesisvial,KRV18}.
These type systems are able to statically capture non-trivial
dynamic properties of terms, particularly {\em normalization},
while at the same time being amenable to elementary proof techniques by induction,
rather than arguments based on reducibility.
Intersection types were originally proposed by
Coppo and Dezani-Ciancaglini~\cite{DBLP:journals/aml/CoppoD78}
to study termination in the $\lambda$-calculus.
They are characterized by the presence of an {\em intersection} type constructor $\typ \cap \typtwo$.
{\em Non-idempotent} intersection type systems are distinguished from their usual idempotent
counterparts by the fact that intersection
is not declared to be idempotent, \ie $\typ$ and $\typ \cap \typ$ are not equivalent types.
Rather, intersection behaves like a multiplicative connective in linear logic.
Arguments to functions are typed many times, typically once
per each time that the argument will be used.
Non-idempotent intersection types were originally formulated by
Gardner~\cite{gardner1994discovering},
and later reintroduced by de Carvalho~\cite{carvalho2007semantiques}.

In this work, we will use a non-idempotent intersection type system
based on system~$\mathcal{W}$~of~\cite{bucciarelli2017non}
(called system~$\mathcal{H}$ in \cite{bucciarelli2014inhabitation}).
Let us recall its definition.
Terms are as usual in the $\lambda$-calculus ($\tm  ::= \var \mid \lam{\var}{\tm} \mid \tm\,\tm$).
Types $\typ,\typtwo,\typthree,\hdots$ are defined by the grammar:
\[
  \begin{array}{rcl@{\hspace{2cm}}rcl}
  \typ  & ::= & \basetyp \OR \mtyp \to \typ
  &
  \mtyp & ::= & [\typ_i]_{i=1}^{n} \HS\text{with $n \geq 0$} \\
  \end{array}
\]
where $\basetyp$ ranges over one of denumerably many {\em base types},
and $\mtyp$ represents a {\em multiset of types}.
Here $[\typ_i]_{i=1}^{n}$ denotes the multiset containing the types $\typ_1,\hdots,\typ_n$
with their respective multiplicities.
A multiset $[\typ_i]_{i=1}^{n}$ intuitively stands for the (non-idempotent) intersection $\typ_1 \cap \hdots \cap \typ_n$.
The {\em sum of multisets} $\mtyp + \mtyptwo$ is defined as their union (adding multiplicities).
A {\em typing context} $\tctx$ is a partial function mapping variables to multisets of types.
The domain of $\tctx$ is the set of variables $\var$ such that $\tctx(\var)$ is defined.
We assume that typing contexts always have {\em finite domain} and hence they may be written as
$\var_1 : \mtyp_1, \hdots, \var_n : \mtyp_n$.
The {\em sum of contexts} $\tctx + \tctxtwo$ is their pointwise sum,
\ie $(\tctx + \tctxtwo)(\var) := \tctx(\var) + \tctxtwo(\var)$ if $\tctx(\var)$ and $\tctxtwo(\var)$ are both defined,
$(\tctx + \tctxtwo)(\var) := \tctx(\var)$ if $\tctxtwo(\var)$ is undefined,
and $(\tctx + \tctxtwo)(\var) := \tctxtwo(\var)$ if $\tctx(\var)$ is undefined.
We write $\tctx +_{i=1}^{n} \tctxtwo_i$ to abbreviate $\tctx + \tctxtwo_1 + \hdots + \tctxtwo_n$.
The {\em disjoint sum of contexts} $\tctx \oplus \tctxtwo$ stands for $\tctx + \tctxtwo$, provided that
the domains of $\tctx$ and $\tctxtwo$ are disjoint.
A {\em typing judgment} is a triple $\tctx \vdash \tm : \typ$,
representing the knowledge that the term $\tm$ has type $\typ$ in the context $\tctx$.
Type assignment rules for system $\mathcal{W}$ are as follows.

\begin{definition}[System $\mathcal{W}$]
\[
  \indrule{var}{
  }{
    \var : [\typ] \vdash \typ
  }
  \indrule{\toI}{
    \tctx \oplus (\var:\mtyp) \vdash \tm : \typ
  }{
    \tctx \vdash \lam{\var}{\tm} : \mtyp \to \typ
  }
  \indrule{\toE}{
    \tctx \vdash \tm : [\typtwo_i]_{i=1}^{n} \to \typ
    \HS
    (\tctxtwo_i \vdash \tmtwo_i : \typtwo_i)_{i=1}^{n}
  }{
    \tctx +_{i=1}^{n} \tctxtwo_i \vdash \tm\,\tmtwo : \typ
  }
\]
\end{definition}
Observe that the \indrulename{\toE} rule has $n + 1$ premises, where $n \geq 0$.
System~$\mathcal{W}$ enjoys various properties, nicely summarized in~\cite{bucciarelli2017non}.

There are two obstacles to adopting system~$\mathcal{W}$ for studying derivation spaces.
The first obstacle is just a matter of presentation---typing derivations use a tree-like notation,
which is cumbersome. One would like to have an alternative presentation based on proof terms.
For example, one would like to write $\var^\typ$ for an application of the \texttt{var} rule,
$\lam{\var}{\tm}$ for an application of the \indrulename{\toI} rule,
and $\tm[\tmtwo_1,\hdots,\tmtwo_n]$ for an application of the
\indrulename{\toE} rule,
so that, for example,
$
  \lam{\var}{\var^{[\alpha,\alpha] \to \beta}[\var^\alpha,\var^\alpha]}
$
represents the following typing derivation:
\[
  \indrule{\toI}{
    \indrule{\toE}{
      \indrule{var}{}{\var : [\alpha,\alpha] \to \beta \vdash \var : [\alpha,\alpha] \to \beta}
      \indrule{var}{}{\var : [\alpha] \vdash \var : \alpha}
      \indrule{var}{}{\var : [\alpha] \vdash \var : \alpha}
      \vspace{.1cm}
    }{
      \var : [[\alpha,\alpha] \to \beta, \alpha,\alpha] \vdash \var\var : \beta
    }
  }{
      \vdash \lam{\var}{\var\var} : [[\alpha,\alpha] \to \beta, \alpha,\alpha] \to \beta
  }
\]
The second obstacle is a major one for our purposes: {\em proof normalization} in this system is not confluent.
The reason is that applications take multiple arguments, and a $\beta$-reduction step must choose a way to
distribute these arguments among the occurrences of the formal parameters.
For instance, the following critical pair cannot be closed:
\[
  \xymatrix@R=.1cm{
  (\lam{\var}{\vartwo^{[\alpha] \to [\alpha] \to \beta}[\var^\alpha][\var^\alpha]})
    [\varthree^{[\gamma] \to \alpha}[\varthree^\gamma],\varthree^{[] \to \alpha}[]]
  \ar[r] \ar[dr]
  &
    \vartwo^{[\alpha] \to [\alpha] \to \beta}[\varthree^{[\gamma] \to \alpha}[\varthree^\gamma]][\varthree^{[] \to \alpha}[]]
  \\
  &
    \vartwo^{[\alpha] \to [\alpha] \to \beta}[\varthree^{[] \to \alpha}[]][\varthree^{[\gamma] \to \alpha}[\varthree^\gamma]]
  }
\]

The remainder of this work is organized as follows:
\begin{itemize}
\item
  In \rcha{preliminaries}, we review some standard notions of order and rewriting theory, as well as some
  basic notions of the $\lambda$-calculus that we will use throughout the work.
\item
  In \rcha{lambdadist}, we introduce a confluent calculus $\lambdadist$ based on system $\mathcal{W}$.
  The desirable properties of system~$\mathcal{W}$ of~\cite{bucciarelli2017non} still hold in $\lambdadist$.
  Moreover, $\lambdadist$ is confluent.
  We impose confluence forcibly, by decorating subtrees with distinct labels, so that
  a $\beta$-reduction step may distribute the arguments in a unique way.
  Derivation spaces in $\lambdadist$ have very regular structure,
  namely they are distributive lattices.
\item
  In \rsec{simulation}, we establish a correspondence between derivation spaces in the
  $\lambda$-calculus and the $\lambdadist$-calculus via simulation theorems,
  which defines a morphism of upper semilattices.
\item
  In \rcha{factorization}, we introduce the notion of a garbage derivation.
  Roughly, a derivation in the $\lambda$-calculus is {\em garbage} if it maps to
  an empty derivation in the $\lambdadist$-calculus.
  This gives rise to an orthogonal notion of {\em garbage-free} derivation.
  The notion of garbage-free derivation is closely related
  with the notions of {\em needed step}~\cite[Section~8.6]{Terese},
  {\em typed occurrence of a redex}~\cite{bucciarelli2017non},
  and {\em external} derivation~\cite{DBLP:conf/ctcs/Mellies97}.
  Using this notion of garbage we prove a {\em factorization theorem}
  reminiscent of Melli\`es'~\cite{DBLP:conf/ctcs/Mellies97}.
  The upper semilattice of derivations of a term in the $\lambda$-calculus
  is factorized using a variant of the Grothendieck construction.
  Every derivation is uniquely decomposed as a garbage-free prefix followed by a garbage suffix.
\item
  In \hyperref[ch:conclusions]{Conclusions}, we end with a discussion of our results.
\end{itemize}
\bigskip
{\bf Note.} Proofs including a \SeeAppendix symbol are spelled out in detail in the appendix.
