\chapter{Residual theory}
\lcha{residual_theory}
In this section we will develop the theory of residuals of the $\lambdadist$-calculus.
We do this as it will help us to prove that derivation spaces of $\lambdadist$
terms have a very simple structure. This together with the fact that
the $\lambdadist$-calculus
can simulate the $\lambda$-calculus in a strong sense (as we will learn in the next chapter)
will make the $\lambdadist$-calculus a very useful tool
to analyze the pure $\lambda$-calculus.


The basic concepts of general rewriting theory were outlined in
the preliminaries (\rsec{rewriting_theory_preliminaries}),
and we will now concentrate in the specifics of our calculus.

Informally, the residual of a step after another is what is left of a step after executing another one; it is a set of steps.

\begin{example} If we consider the term of the pure lambda calculus:
\[(\lam{\var}{\vartwo})((\lam{\varthree}{\varthree}) w)\]
Then there are two steps we can perform (we shall call them $\redex$ and $\redextwo$).
\begin{align*}
  \redex &:
    (\lam{\var}{\vartwo})((\lam{\varthree}{\varthree}) w) \to y \\
  \redextwo &:
    (\lam{\var}{\vartwo})((\lam{\varthree}{\varthree}) w) \to (\lam{\var}{\vartwo}) w
\end{align*}
The residual of $\redex$ after $\redextwo$ is the step $(\lam{\var}{\vartwo}) w \to w$.
On the other hand, the residual of $\redextwo$ after $\redex$ is the empty set:
$\redextwo$ was erased by $\redex$.
\end{example}

It is interesting to develop a theory of residuals for $\lambdadist$ because
residuals allow us to trace a redex through the reduction of a term.
Given that the purpose of $\lambdadist$ is to be able to track
how resources are used in $\lambda$-terms, it is crucial to
be able to learn how these resources interact during a reduction.

As it turns out, the theory of residuals of $\lambdadist$ will prove to be powerful
enough to represent meaningful information, but simple enough to have a
comprehensible structure.


\begin{definition}
If $\redex : \tm \todistl{\lab} \tm'$ is a step in the distributive lambda
calculus, then:
\[
  \begin{array}{rcll}
    \src(\redex) & \eqdef & \tm & \text{ is the \defn{source} of $\redex$} \\
    \tgt(\redex) & \eqdef & \tm' & \text{ is the \defn{target} of $\redex$} \\
    \name(\redex) & \eqdef & \lab & \text{ is the \defn{name} of $\redex$}
  \end{array}
\]
Two steps $\redex$ and $\redextwo$
are \defn{coinitial} if $\src(\redex) = \src(\redextwo)$
and \defn{cofinal} if $\tgt(\redex) = \tgt(\redextwo)$.
\end{definition}

\begin{definition}[Residuals in the distributive lambda-calculus]
\ldef{residuals}
Given coinitial steps $\redex, \redextwo$,
the set $\redex/\redextwo$ of \defn{residuals of $\redex$ after $\redextwo$} is defined as follows:
\[
  \redex/\redextwo = \set{\redex' \ST \src(\redex') = \tgt(\redextwo) \text{ and } \name(\redex) = \name(\redex')}
\]
\end{definition}

\begin{remark}
Recall that the name of a step is the label that decorates the lambda reduced by the step,
and that in correct terms all lambdas have pairwise distinct labels.
Given that our calculus has no duplication or erasure (as per the following lemma),
names of steps will be useful to name reductions.
\end{remark}


\begin{lemma}[Cardinality of the set of residuals]
\llem{cardinality_of_the_set_of_residuals}
\[
  \#(\redex/\redextwo) = \begin{cases}
                         0 & \text{ if $\redex = \redextwo$} \\
                         1 & \text{ otherwise}
                         \end{cases}
\]
\end{lemma}
\begin{proof}
Recall that, by definition, lambdas in a correct term have pairwise distinct labels.
Consider first the case when $\redex = \redextwo$.
Then
$
  \redex = \redextwo : \conof{(\lamp{\lab}{\var}{\tm})\ls{\tmtwo}} \todistl{\lab} \conof{\subs{\tm}{\var}{\ls{\tmtwo}}}
$.
There is only one lambda decorated with $\lab$ in the source,
so there are no lambdas decorated with $\lab$ in the target.
Hence $\redex/\redextwo = \emptyset$.

On the other hand if, $\redex \neq \redextwo$,
by the Strong Permutation property (\rprop{strong_permutation})
there exists a step $\redex' \in \redex/\redextwo$
with the same name as $\redex$.
There are no other lambdas decorated with $\lab$ in the target.
Hence $\redex/\redextwo = \set{\redex'}$.
\end{proof}


\section{Orthogonality of $\lambdadist$}
\input{04-01-orthogonality}

\section{Names and labels}
\input{04-02-names}

\section{Stability (and creation)}
\input{04-03-creation-and-stability}

\section{Lattices and derivation spaces}
\input{04-04-lattices}




