What follows are two results that describe what happens when one executes a garbage step or garbage derivation.
In short, if a garbage derivation creates or duplicates a step $\redex$,
then $\redex$ is garbage.
This makes sense, as one should not need to execute garbage in order to create a ``needed'' step.

\begin{lemma}[Garbage only creates garbage]
\llem{garbage_only_creates_garbage}
Let $\redex$ and $\redextwo$ be composable steps in the $\lambda$-calculus,
and let $\tm' \refines \src(\redex)$.
If $\redex$ creates $\redextwo$ and $\redex$ is $\tm'$-garbage,
then $\redextwo$ is $(\tm'/\redex)$-garbage.
\end{lemma}
\begin{proof} According to L\'evy~\cite{Tesis:Levy:1978},
there are three creation cases in the $\lambda$-calculus.
We need to consider those three cases separately. \SeeAppendixRef{garbage_only_creates_garbage_proof}
\end{proof}

\begin{lemma}[Garbage only duplicates garbage]
\llem{garbage_only_duplicates_garbage}
Let $\redex$ and $\redextwo$ be coinitial steps in the $\lambda$-calculus
and let $\tm' \refines \src(\redex)$.
If $\redex$ duplicates $\redextwo$, \ie $\#(\redextwo/\redex) > 1$,
and $\redex$ is $\tm'$-garbage,
then $\redextwo$ is $(\tm'/\redex)$-garbage.
\end{lemma}
\begin{proof}
\SeeAppendixRef{garbage_only_duplicates_garbage_proof} By inspecting how $\redex$ looks like given that it duplicates
$\redextwo$.
\end{proof}

We now prove two propositions that give different characterizations for both garbage and
garbage-free derivations. Finally, we give some general and useful properties of the
sieving operation.

\begin{proposition}[Characterization of garbage]
\lprop{characterization_of_garbage}
Let $\redseq : \tm \rtobeta \tmtwo$ and $\tm' \refines \tm$.
Then the following are equivalent:
\begin{enumerate}
\item $\redseq \sieve \tm' = \emptyDerivation$.
\item There are no coarse steps for $(\redseq,\tm')$.
\item The derivation $\redseq$ is $\tm'$-garbage.
\end{enumerate}
\end{proposition}
\begin{proof}
It is immediate to check that items 1 and 2 are equivalent, by definition of sieving,
so we focus on $2 \implies 3$ and $3 \implies 2$. \SeeAppendixRef{characterization_of_garbage_proof}
\end{proof}

\begin{proposition}[Characterization of garbage-free derivations]
\lprop{characterization_of_garbage_free_derivations}
Let $\redseq : \tm \rtobeta \tmtwo$ and $\tm' \refines \tm$.
Then the following are equivalent:
\begin{enumerate}
\item $\redseq$ is $\tm'$-garbage-free,
\item $\redseq \permeq \redseq \sieve \tm'$,
\item $\redseq \permeq \redseqtwo \sieve \tm'$ for some derivation $\redseqtwo$.
\end{enumerate}
\end{proposition}
\begin{proof}
We prove each direction separately. \SeeAppendixRef{characterization_of_garbage_free_derivations_proof}
\end{proof}


\begin{proposition}[Properties of sieving]
\lprop{properties_of_sieving}
Let $\tm' \refines \tm$ and $\redseq : \tm \tobeta^* \tmtwo$. Then:
\begin{enumerate}
\item $\redseq \sieve \tm'$ is $\tm'$-garbage-free and $\redseq \sieve \tm' \permle \redseq$.
\item $\redseq/(\redseq \sieve \tm')$ is $(\tm'/(\redseq \sieve \tm'))$-garbage.
\item $\redseq$ is $\tm'$-garbage if and only if $\redseq \sieve \tm' = \emptyDerivation$.
\item $\redseq$ is $\tm'$-garbage-free if and only if $\redseq \sieve \tm' \permeq \redseq$.
\end{enumerate}
\end{proposition}
\begin{proof}
We use some technical lemmas spelled out in the appendix (\SeeAppendixRef{properties_of_sieving_proof}).
\begin{enumerate}
\item  Note that $\redseq \sieve \tm'$ is $\tm'$-garbage-free by
  \rprop{characterization_of_garbage_free_derivations}.
  Moreover, $\redseq \sieve \tm' \permle \redseq$ by~\rlem{sieve_is_prefix}.
\item This is precisely~\rlem{projection_after_sieving_is_garbage}.
\item An immediate consequence of~\rprop{characterization_of_garbage}.
\item An immediate consequence of~\rprop{characterization_of_garbage_free_derivations}.
\end{enumerate}
\end{proof}

